### 2014-10-27 motivation.tex
motivation
Edits

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{Figures/SystemModel}
\caption{Quelea system model.}
\label{fig:sysmod}
\end{figure}

@@

When \emph{foo} is invoked on a object $x$ with arguments \emph{arg}$_1$ at
replica 1, it simply \emph{reduces} over the current set of
effects at that replica on that object ($w^x_1$ and $w^x_2$), produces a result
$v1$ that is sent back to the client, and emits a \emph{single} new effect
$w^x_4$ that is appended to the state $x$ at replica 1. Thus, every operation
is evaluated over a \emph{snapshot} of the state of the object on which it is
invoked. In this case, we say that the effect $w^x_4$ \emph{witnesses} $w^x_1$
and $w^x_2$. Alternatively, $w^x_1$ and $w^x_2$ are \emph{visible} to $w^x_4$,
written logically as $\vis{w^x_1}{w^x_4} \wedge \vis{w^x_2}{w^x_4}$, where
$\visZ$ is the visibility relation between effects. Visibility is an
irreflexive (an operation cannot see its own effect) and asymmetric relation,
and only relates effects produced by operations on the same object.
Executing a read-only operation is similar except
that no new effects are produced.

@@

This model admits all the inconsistencies associated with eventual
consistency. The goal of this work is to identify the precise consistency
level for each operation such that application-level constraints are not
violated. In the next section, we will concretely describe the challenges
associated with constructing a consistent bank account on top of an
eventually consistent data stores. Subsequently, we will illustrate how our
contract and specification language, armed with the primitive relations
$\visZ$, $\soZ$, $\sameobjZ$ and $\operZ$, mitigates these challenges.

\subsection{Eventually Consistent Bank Account}

Consider how we might implement a highly available bank account on top of an
eventually consistent data store, with the \emph{integrity} constraint that
the balance should never drop below 0. Let us assume that the bank account
object provides three operations: \cf{deposit}, \cf{withdraw} and
\cf{getBalance}, with the assumption that the withdraw fails if the account
has insufficient balance. Our goal is to choose the correct consistency
level for each of these operations such that (1) the balance never drops
below zero and (2) the \cf{getBalance} operation never incorrectly show a
negative balance. To be able to do this, we need to consider the anomalies
that could arise under eventual consistency.

@@

%% \begin{figure}
%% \centering
%% \includegraphics[width=\columnwidth]{Graphs/BankAccount_Th_vs_Lat}
%% \caption{Bank Account: Throughput vs Latency. Quelea is our implementation.
%% Strong and Quelea are correct. High. Avail. and causal are incorrect (withdraw
%% not strong). 20\% withdraw, 20\% deposit, 60\% getBalance.}
%% \label{grf:BankAcc_Th_vs_Lat}
%% \end{figure}

@@

%%  a strongly consistent operation is
%% significantly expensive (Figure~\ref{grf:BankAcc_Th_vs_Lat} \KC{Should I
%%   include the numbers here? If yes, I'll expand the text.}), both in terms
%% of availability and monetary overheads~\cite{DynamoDB,Pileus}, we ideally
%% would not want to request strong consistency, unless there are no
%% alternatives available.

\subsection{Contracts}

@@

In our running example, it is clear that in order to preserve the integrity
constraint, the \cf{withdraw} operation must be strongly consistent.  That
is, given two \cf{withdraw} operations $a$ and $b$, either $a$ is visible to
$b$ or vice versa. Utilizing the primitive relations described in
Section~\ref{sec:sysmod}, we express this application-level consistency
requirement as a contract or specification over \cf{withdraw}:
\begin{smathpar}
\begin{array}{l}
\forall (a : \rcf{withdraw}).~\sameobj{a}{\cureff} \Rightarrow a = \cureff \vee \vis{a}{\cureff} \vee \vis{\cureff}{a}
\end{array}
\end{smathpar}
Here, $\cureff$ stands for the effect emitted by the \cf{withdraw}
operation. The syntax $a : \rcf{withdraw}$ states that $a$ is an effect
which was emitted by a \cf{withdraw} operation i.e.,
$\oper{a}{\rcf{withdraw}}$ holds.  The contract specifies that if the
current operation emits an effect $\cureff$, then for any operation $a$
which was emitted by a \cf{withdraw} operation, it is the case that $a =
\cureff$ or $a$ is visible to $\cureff$, or vice versa. Any execution on a
bank account object that preserves the above contract for a \cf{withdraw}
operation is said to be derived from a correct implementation of
\cf{withdraw}.

For \cf{getBalance}, we construct the following contract:
\begin{smathpar}
\begin{array}{l}
\forall (a:\rcf{deposit}), (b:\rcf{withdraw}), (c: \rcf{deposit} \vee \rcf{withdraw}). \\
\qquad \vis{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff} \\
\qquad \wedge~ (\soZ \cap \sameobjZ) (c,\cureff) \Rightarrow \vis{c}{\cureff}
\end{array}
\end{smathpar}

If a \cf{withdraw} $b$ is visible to \cf{getBalance}, then all \cf{deposit}
operations $a$ visible to the \cf{withdraw} should also be visible to
\cf{getBalance}. This prevents negative balance anomalies. The expression $c
: \rcf{deposit} \vee \rcf{withdraw}$ states that $c$ is an effect that was
emitted either by a \cf{deposit} or a \cf{withdraw} operation.  Additonally,
our specification language provides relational operators to compose
relations. Thus, the syntax $(R_1 \cap R_2)(a,b)$ is equivalent to $R_1(a,b)
\wedge R_2(a,b)$. The last line of the above contract says that if a
\cf{deposit} or a \cf{withdraw} precedes a \cf{getBalance} operation in
session order, and is applied on the same object as the \cf{getBalance}
operation, then it must be the case that the \cf{getBalance} operation
witnesses the effects of the preceding operations.  Since there are no
restrictions on when and how a \cf{deposit} operation can execute, its
contract is simply $\true$.

@@

We devise a static approach instead.  Contracts are analyzed with the help
of a theorem prover, and statically mapped to a particular store-level
consistency property that the prover guarantees preserves contract
semantics. The key idea is that the system implementor axiomatically
describes the store semantics in the \emph{same} contract language as the
one used to express the contract on the operation.

>>>>

\begin{figure}
\centering
\includegraphics[width=0.75\columnwidth]{Figures/SystemModel}
\caption{Quelea system model.}
\label{fig:sysmod}
\end{figure}

@@

When \emph{foo} is invoked on a object $x$ with arguments \emph{arg}$_1$ at
replica 1, it simply \emph{reduces} over the current set of effects at that
replica on that object ($w^x_1$ and $w^x_2$), produces a result $v1$ that is
sent back to the client, and emits a \emph{single} new effect $w^x_4$ that is
appended to the state of $x$ at replica 1. Thus, every operation is evaluated
over a \emph{snapshot} of the state of the object on which it is invoked. In
this case, we say that the effect $w^x_4$ \emph{witnesses} $w^x_1$ and $w^x_2$.
Alternatively, $w^x_1$ and $w^x_2$ are \emph{visible} to $w^x_4$, written
logically as $\vis{w^x_1}{w^x_4} \wedge \vis{w^x_2}{w^x_4}$, where $\visZ$ is
the visibility relation between effects. Visibility is an irreflexive (an
operation cannot see its own effect) and asymmetric relation, and only relates
effects produced by operations on the same object. Executing a read-only
operation is similar except that no new effects are produced.

@@

This model admits all the inconsistencies associated with eventual consistency.
The goal of this work is to identify the precise consistency level for each
operation such that application-level constraints are not violated. In the next
section, we will concretely describe the challenges associated with
constructing a consistent bank account on top of an eventually consistent data
store. Subsequently, we will illustrate how our contract and specification
language, armed with the primitive relations $\visZ$, $\soZ$, $\sameobjZ$ and
$\operZ$, mitigates these challenges.

\subsection{Example}

Consider how we might implement a highly available bank account on top of an
eventually consistent data store, with the \emph{integrity} constraint that the
balance should never drop below 0. Let us assume that the bank account object
provides three operations: \cf{deposit}, \cf{withdraw} and \cf{getBalance},
with the assumption that the withdraw fails if the account has insufficient
balance. Our goal is to choose the correct consistency level for each of these
operations such that (1) the balance never drops below zero and (2) the
\cf{getBalance} operation never incorrectly returns a negative balance. To be
able to do this, we need to consider the anomalies that could arise under
eventual consistency.

@@

@@

\subsection{Specifications and Contracts}

@@

In our running example, it is clear that in order to preserve the integrity
constraint, the \cf{withdraw} operation must be strongly consistent.  That is,
given two \cf{withdraw} operations $a$ and $b$, either $a$ is visible to $b$ or
vice versa. Utilizing the primitive relations described in
Section~\ref{sec:sysmod}, we express this application-level consistency
requirement as a contract or specification over \cf{withdraw}:

\begin{smathpar}
\begin{array}{l}
\forall (a : \rcf{withdraw}).~\sameobj{a}{\cureff} \Rightarrow a = \cureff \vee \vis{a}{\cureff} \vee \vis{\cureff}{a}
\end{array}
\end{smathpar}

Here, $\cureff$ stands for the effect emitted by the \cf{withdraw} operation.
The syntax $a:\rcf{withdraw}$ states that $a$ is an effect which was emitted
by a \cf{withdraw} operation i.e., $\oper{a}{\rcf{withdraw}}$ holds.  The
contract specifies that if the current operation emits an effect $\cureff$,
then for any operation $a$ which was emitted by a \cf{withdraw} operation, it
is the case that $a = \cureff$ or $a$ is visible to $\cureff$, or vice versa.
Any execution on a bank account object that preserves the above contract for a
\cf{withdraw} operation is said to be derived from a correct implementation of
\cf{withdraw}.

For \cf{getBalance}, we construct the following contract:

\begin{smathpar}
\begin{array}{l}
\forall (a:\rcf{deposit}), (b:\rcf{withdraw}), (c: \rcf{deposit} \vee \rcf{withdraw}). \\
\qquad \vis{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff} \\
\qquad \wedge~ (\soZ \cap \sameobjZ) (c,\cureff) \Rightarrow \vis{c}{\cureff}
\end{array}
\end{smathpar}

The expression $c:\rcf{deposit} \vee \rcf{withdraw}$ states that $c$ is an
effect that was emitted either by a \cf{deposit} or a \cf{withdraw} operation.
If a \cf{withdraw} $b$ is visible to \cf{getBalance}, then all \cf{deposit}
operations $a$ visible to the \cf{withdraw} should also be visible to
\cf{getBalance}. This prevents negative balance anomalies.

Let us consider the last line in the above contract. Our specification language
provides operators to for composing relations. Thus, the syntax $(R_1 \cap
R_2)(a,b)$ is equivalent to $R_1(a,b) \wedge R_2(a,b)$. The last line of the
above contract says that if a \cf{deposit} or a \cf{withdraw} precedes a
\cf{getBalance} operation in session order, and is applied on the same object
as the \cf{getBalance} operation, then it must be the case that the
\cf{getBalance} operation witnesses the effects of the preceding operations.

Since there are no restrictions on when and how a \cf{deposit} operation can
execute, its contract is simply $\true$.

@@

We devise a static approach instead. Contracts are analyzed with the help of a
theorem prover, and statically mapped to a particular store-level consistency
property that the prover guarantees preserves contract semantics. We call this
procedure \emph{contract classification}. Given the variety and complexity of
store level consistency properties, the idea is that the system implementor
parameterizes the classification procedure by describing the store semantics in
the \emph{same} contract language as the one used to express the contract on
the operations. In the next section, we describe the contract language in
detail and describe the classification procedure for a particular store
semantics.
